Random Forests", OAs

### 2023-11-21 src/la_RF.Rmd
CAST

@@

  set.seed(71)
  indices <- CreateSpacetimeFolds(df %>% filter(RGN21CD==i), 
                                spacevar = "ladcd",
                                timevar = "year", 
                                k = 10)

  set.seed(71)
  model.all <- train(n ~ area + dist + dist.retail + London + south + 
                     n.slag + n.lag + n.l.slag + n.nearest.city + year,
                   data = df %>% filter(RGN21CD==i), 
                   trControl = tc,
                   method = "ranger",
                   na.action = na.omit,
                   #preProc = c("center", "scale"),
                   importance = "impurity")
  assign(paste0("model.all",as.character(i)), model.all)
}

@@

path.out <- paste0(path, "/outputs/rf/heatmap_LA.png")
rf.year.all.metrics %>% 
  as.data.frame() %>%
  rownames_to_column(var = "metrics") %>% 
  pivot_longer(!metrics, names_to = "train.test") %>% 
  pivot_wider(names_from = metrics, values_from = value) %>% 
  separate(train.test, c("train", "test"), remove = F) %>% 
  left_join(lookup.region, by = c("train" = "RGN21CD")) %>% 
  rename(train.region=RGN21NM) %>% 
  left_join(lookup.region, by = c("test" = "RGN21CD")) %>% 
  rename(test.region=RGN21NM) %>% 
  arrange(Rsquared) %>% 
  dplyr::select(train.region, test.region, Rsquared) %>% 
  ggplot(aes(train.region, test.region, fill= Rsquared)) + 
  geom_tile() +
  coord_fixed() +
  labs(y = "", x = "") +
  scale_fill_gradientn(colours = terrain.colors(10), trans = 'reverse') +
  theme(#legend.position = "none",
        axis.text.x = element_text(angle = 45, hjust=1),
        panel.border = element_blank(), 
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_rect(fill = "white"))
ggsave(path.out)

>>>>

## unregister_dopar function

This is helpful when the parallel loop collapses.

```{r}
unregister_dopar <- function() {
  env <- foreach:::.foreachGlobals
  rm(list=ls(name=env), pos=env)
}
unregister_dopar()
```

@@

  set.seed(71)
  indices <- CreateSpacetimeFolds(df %>% filter(RGN21CD==i), 
                                #spacevar = "ladcd",
                                timevar = "year", 
                                k = 10)

  model.all <- train(n ~ area + dist + dist.retail + London + south + 
                     n.slag + n.lag + n.l.slag + n.nearest.city + year,
                   data = df %>% filter(RGN21CD==i), 
                   trControl = tc,
                   method = "ranger",
                   na.action = na.omit,
                   #preProc = c("center", "scale"),
                   importance = "impurity")
  assign(paste0("model.all",as.character(i)), model.all)
}

@@

path.out <- paste0(path, "/outputs/rf/heatmap_LA.png")
rf.year.all.metrics %>% 
  as.data.frame() %>%
  rownames_to_column(var = "metrics") %>% 
  pivot_longer(!metrics, names_to = "train.test") %>% 
  pivot_wider(names_from = metrics, values_from = value) %>% 
  separate(train.test, c("train", "test"), remove = F) %>% 
  left_join(lookup.region, by = c("train" = "RGN21CD")) %>% 
  rename(train.region=RGN21NM) %>% 
  left_join(lookup.region, by = c("test" = "RGN21CD")) %>% 
  rename(test.region=RGN21NM) %>% 
  arrange(Rsquared) %>% 
  dplyr::select(train.region, test.region, Rsquared) %>% 
  ggplot(aes(train.region, test.region, fill= Rsquared)) + 
  geom_tile() +
  coord_fixed() +
  labs(y = "", x = "") + #y: test.region, x: train.region
  scale_fill_gradientn(colours = terrain.colors(10), trans = 'reverse') +
  theme(#legend.position = "none",
        axis.text.x = element_text(angle = 45, hjust=1),
        panel.border = element_blank(), 
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_rect(fill = "white"))
ggsave(path.out)

### 2023-11-21 src/oa_RF.Rmd
CAST

## spatial data

@@

@@

```{r}
oa.sf <- st_as_sf(oa.gb)

oa.full <- replicate(17, oa.sf %>% as_tibble(), simplify = FALSE) %>%
  bind_rows(.id = 'id_') %>% 
  rename(year = id_) %>% 
  mutate(year = as.numeric(year) + 1995) %>% 
  dplyr::select(id, year)

@@

```{r}
# oa centroids
# oa.gb.c <- (rgeos::gCentroid(oa.gb, byid=TRUE))@coords %>% as.data.frame()
# oa.gb.c$id <- row.names(oa.gb.c)  #as_tibble(rownames = NA) %>% 
#   mutate(id = row.names())
oa.gb.c <- gCentroid(oa.gb, byid=TRUE)@coords %>% 
  as.data.frame() %>% 
  add_rownames(var = "id")

@@

# retail centres
geo.path <- paste0(path, "/data/raw/Retail_Boundaries_UK.gpkg")
retail <-readOGR(geo.path) 
#retail.major.cetres <- subset(retail, retail$Classification == "Major Town Centre") #%>% 
retail.major.cetres.help <- gCentroid(retail, byid=TRUE)@coords %>% 
  as.data.frame() %>% 
  add_rownames(var = "id") #%>%

retail.major.cetres <- retail.major.cetres.help %>% 
  st_as_sf(coords = c("x", "y"), crs = 27700) %>%
  st_transform(4326) %>%
  st_coordinates() %>%
  as_tibble() %>% 
  bind_cols(retail.major.cetres.help$id) %>% 
  rename(id = '...3') %>% 
  left_join(retail@data %>% mutate(id = rownames(retail@data)), by = 'id')
```

@@

```{r}
dist <- distm(cbind(oa.gb.c$x, oa.gb.c$y), cbind(cities$long, cities$lat), fun=distHaversine) 
dist <- round((dist/1000),2) %>% 
  as_tibble()

@@

# calculate area
oa.gb$area <- area(oa.gb) #/ 1000000

# Join with distance and area
df <- df %>% left_join(dist, by = c("id" = "oa11cd")) %>% 
  rename(oa11cd=id) %>% 
  left_join(oa.gb@data, by = c("oa11cd" = "id")) %>% 
  relocate(area, .after = n) %>% 
  mutate(area = area / 1000000,
         density = n / area) %>% 
  relocate(density, .after = area)

@@

```{r}
dist.retail <- distm(cbind(oa.gb.c$x, oa.gb.c$y), cbind(retail.major.cetres$X, retail.major.cetres$Y), fun=distHaversine) 
dist.retail <- round((dist.retail/1000),2) %>% 
  as_tibble()

@@

sf_use_s2(FALSE)
help1 <- st_filter( oa.sf, cities.sf) %>% as_tibble %>% dplyr::select(id) %>% 
  mutate(cities = c("Leicester",
                    "Bristol",
                    "Manchester",
                    "Liverpool",
                    "Sheffield",
                    "Birmingham",
                    "Leeds",
                    "London",
                    "Edinburgh",
                    "Glasgow"))

@@

#####save this#####
sf_use_s2(FALSE)
oa.c <- st_centroid(oa.sf)
kn <- knearneigh(oa.c, k = 5)
knn <- knn2nb(kn, row.names = NULL, sym = FALSE)
knn.l <- nb2listw(knn)
##################

>>>>

## spatial data OLD

@@

## spatial data NEW

```{r}
library(sf)

# Get OA for England and Wales
path.geo <- paste0(path, "/data/raw/Output_Areas__December_2011__Boundaries_EW_BGC.geojson")
oa.ew <- st_read(path.geo)

# Spatial transformations
oa.ew <- st_transform(oa.ew, 4326)  # EPSG code for WGS84

# Rename columns and keep only necessary columns
oa.ew <- oa.ew %>% dplyr::select(OA11CD, Shape__Area) %>% 
  rename(id = OA11CD,
         area = Shape__Area)
# oa.ew <- oa.ew[, c('OA11CD', 'Shape__Area')]
# colnames(oa.ew) <- c('id', 'area')

# Check the data structure
print(summary(oa.ew))

# Get OA for Scotland
path.geo.sc <- paste0(path, "/data/raw/output-area-2011-mhw")
oa.sc <- st_read(dsn = path.geo.sc, layer = "OutputArea2011_MHW")

# Spatial transformations
oa.sc <- st_transform(oa.sc, 4326)  # EPSG code for WGS84

# Rename columns and keep only necessary columns
oa.sc <- oa.sc %>% dplyr::select(code, SHAPE_1_Ar) %>% 
  rename(id = code,
         area = SHAPE_1_Ar)
# oa.sc <- oa.sc[, c('code', 'SHAPE_1_Ar')]
# colnames(oa.sc) <- c('id', 'area')

# Check the data structure
print(summary(oa.sc))

# Build OA for GB
oa.gb <- rbind(oa.ew, oa.sc)
rm(oa.ew, oa.sc)

# Get GB
path.gb <- "/hdd/internet_archive/archive/gis/Countries_December_2014_Full_Clipped_Boundaries_in_Great_Britain.shp"
gb <- st_read(path.gb)

# Spatial transformations
gb <- st_transform(gb, 4326)  # EPSG code for WGS84
```

@@

```{r}
#oa.sf <- st_as_sf(oa.gb)

oa.full <- replicate(17, oa.gb %>% as_tibble(), simplify = FALSE) %>%
  bind_rows(.id = 'id_') %>% 
  rename(year = id_) %>% 
  mutate(year = as.numeric(year) + 1995) %>% 
  dplyr::select(id, year)

@@

```{r}
# oa centroids
# oa.gb.c <- (rgeos::gCentroid(oa.gb, byid=TRUE))@coords %>% as.data.frame()
# oa.gb.c$id <- row.names(oa.gb.c)  #as_tibble(rownames = NA) %>% 
#   mutate(id = row.names())

# oa.gb.c <- gCentroid(oa.gb, byid=TRUE)@coords %>% 
#   as.data.frame() %>% 
#   add_rownames(var = "id")
sf_use_s2(FALSE)
oa.gb.c <- st_centroid(oa.gb) %>%
  st_coordinates() %>%
  as.data.frame() %>%
  #add_rownames(var = "id") %>% 
  bind_cols(oa.gb %>% dplyr::select(id))

@@

# retail centres
geo.path <- paste0(path, "/data/raw/Retail_Boundaries_UK.gpkg")
#retail <-readOGR(geo.path) 
retail <- st_read(geo.path)

#retail.major.cetres <- subset(retail, retail$Classification == "Major Town Centre") #%>% 
# retail.major.cetres.help <- gCentroid(retail, byid=TRUE)@coords %>% 
#   as.data.frame() %>% 
#   add_rownames(var = "id") #%>%
retail.major.cetres.help <- st_centroid(retail) %>% 
  st_transform(4326) %>%
  st_coordinates() %>% 
  as.data.frame() %>% 
  add_rownames(var = "id")

# retail.major.cetres <- retail.major.cetres.help %>% 
#   st_as_sf(coords = c("x", "y"), crs = 27700) %>%
#   st_transform(4326) %>%
#   st_coordinates() %>%
#   as_tibble() %>% 
#   bind_cols(retail.major.cetres.help$id) %>% 
#   rename(id = '...3') %>% 
#   left_join(retail@data %>% mutate(id = rownames(retail@data)), by = 'id')
retail.major.cetres <- retail.major.cetres.help %>% 
  # as_tibble() %>% 
  # bind_cols(retail.major.cetres.help$id) %>% 
  # #rename(id = '...3') %>% 
  left_join(retail %>% rownames_to_column(var = "id"), by = 'id')
```

@@

```{r}
dist <- distm(cbind(oa.gb.c$Y, oa.gb.c$Y), cbind(cities$long, cities$lat), fun=distHaversine) 
dist <- round((dist/1000),2) %>% 
  as_tibble()

@@

# calculate area
oa.gb$area <- st_area(oa.gb) #/ 1000000

# Join with distance and area
df <- df %>% left_join(dist, by = c("id" = "oa11cd")) %>% 
  rename(oa11cd=id) %>% 
  left_join(oa.gb, by = c("oa11cd" = "id")) %>% 
  relocate(area, .after = n) %>% 
  mutate(area = area / 1000000,
         density = n / area) %>% 
  relocate(density, .after = area) %>% 
  dplyr::select(-geometry)

@@

```{r}
dist.retail <- distm(cbind(oa.gb.c$X, oa.gb.c$Y), cbind(retail.major.cetres$X, retail.major.cetres$Y), fun=distHaversine) 
dist.retail <- round((dist.retail/1000),2) %>% 
  as_tibble()

@@

sf_use_s2(FALSE)
help1 <- st_filter( oa.gb, cities.sf) %>% as_tibble %>% dplyr::select(id) %>% 
  mutate(cities = c("Leicester",
                    "Bristol",
                    "Manchester",
                    "Liverpool",
                    "Sheffield",
                    "Birmingham",
                    "Leeds",
                    "London",
                    "Edinburgh",
                    "Glasgow"))

@@

#####save this#####
sf_use_s2(FALSE)
oa.c <- st_centroid(oa.gb)
kn <- knearneigh(oa.c, k = 5)
knn <- knn2nb(kn, row.names = NULL, sym = FALSE)
knn.l <- nb2listw(knn)
##################
